# Multi-format projects

One of the main benefits of R Markdown can create multiple output formats from a single source document. This book example is available in all three formats, with PDF, ebook and online versions all being available.

Project should therefore ideally be designed to be flexible for the multiple outputs. However, users will often find themselves wanting to fine-tune the output, and doing so will often require

This chapter aims to provide several examples on how a document can be customised with output specific features, whilst also retaining the ability to export documents to multiple formats.

## Output-specific functions

HTML documents have the capacity to contain. The webshot package is available to take screenshots of HTML elements for non-interactive output formats (ebooks, PDF).

```{r eval=FALSE, echo = TRUE}
library(ggplot2)
library(plotly)
p <- ggplot(data = diamonds, aes(x = cut, fill = clarity)) +
            geom_bar(position = "dodge")
ggplotly(p)
```

## Designing output

If we wish to design our own format-specific function we can use the functions `knitr::is_latex_output()` and `knitr::is_html_output`. These function will return a TRUE/FALSE action

```{r, eval = FALSE}
knitr::is_html_output()
```

This can be used to control the output of statements.

For example, you may want to have interactive tables using `DT:datatable` in the HTML output but print static versions in the PDF.





