\mainmatter

# Installation

To use R Markdown, you have to install R [@R-base] and the R package **rmarkdown** [@R-rmarkdown]. We also recommend that you install the RStudio IDE (https://www.rstudio.com). RStudio is not required, but it will make it easier for an average user to work with R Markdown because of the editor support. If choose not to use the RStudio IDE, you will need a separate installation of Pandoc\index{Pandoc} (https://pandoc.org).

```{r eval=FALSE}
# install the rmarkdown package from CRAN in R
install.packages('rmarkdown')

# or install from GitHub if you want to test the development version
if (!requireNamespace("remotes")) install.packages('remotes')
remotes::install_github('rstudio/rmarkdown')
```

## Install LaTeX for PDF reports {#install-latex}

If you would like to create PDF documents from R Markdown, you will need to have a LaTeX distribution installed. Although there are several traditional options including MiKTeX, MacTeX, and TeX Live, we recommend that R Markdown users install [TinyTeX](https://yihui.name/tinytex/).

TinyTeX is a custom LaTeX distribution based on TeX Live that is small in size (150Mb on macOS/Linux and 220Mb on Windows when installed) but functions well in most cases, especially for R users. Installing or running TinyTeX does not require sysadmin privileges. You can install TinyTeX with the R package **tinytex** [@R-tinytex]:

```{r eval=FALSE}
tinytex::install_tinytex()
# to uninstall TinyTeX, run tinytex::uninstall_tinytex() 
```

Please note that "**tinytex**" refers to the R package, and "TinyTeX" refers to the LaTeX distribution. There are two advantages of using TinyTeX: 

1. TinyTeX is lightweight (compared to other LaTeX distributions), cross-platform, and portable. For example, you can store a copy of TinyTeX on your USB drive or other portable devices, and use it on other computers with the same operating system. 

1. When R Markdown is converted to PDF, Pandoc converts Markdown to an intermediate LaTeX document first. The R package **tinytex** has provided helper functions to compile LaTeX documents to PDF (the main function is `tinytex::latexmk()`). If you use TinyTeX and certain LaTeX packages are required but not installed, **tinytex** will try to automatically install them for you. It will also try to compile the LaTeX file for a sufficient number of times to make sure all cross-references are resolved.

If you are interested in the technical details, you may check out the article @tinytex2019 and the FAQ page at https://yihui.name/tinytex/faq/.
