# Chunk Hooks {#chunk-hooks}

A chunk hook is a function that is triggered by a chunk option when the value of this chunk option is not `NULL`. Chunk hooks provide a way for you to execute additional tasks beyond running the code in a chunk.

You can use chunk hooks purely for their side-effects (e.g., only printing out certain information to the console), or for their returned values, which will be written to the output document if the value is a character value.

Like output hooks (see Chapter \@ref(output-hooks)), chunk hooks are also registered via the object `knitr::knit_hooks`. Please note that the names of output hooks are reserved by **knitr**, so you must not use these names for your custom chunk hooks:

```{r}
names(knitr:::.default.hooks)
```

A chunk hook is associated with a chunk option of the same name. For example, you can register a chunk hook with the name `greet`:

```{r}
knitr::knit_hooks$set(greet = function(before) {
  if (before) "Hello!" else "Bye!"
})
```

We will explain the arguments of the hook function in a moment. Now we set the chunk option `greet = TRUE` for the chunk below: 

````md
```{r, greet=TRUE}`r ''`
1 + 1
```
````

And you will see that "Hello!" appears before the chunk, and "Bye!" appears after the chunk in the output below (which is because they are character values):

::: {.infobox .tip data-latex="{tip}"}
```{r, greet=TRUE}
1 + 1
```
:::

A chunk hook function can possibly take four arguments: `before`, `options`, `envir`, and `name`. In other words, it can be of this form:

```r
function(before, options, envir, name) {

}
```

All four arguments are optional. You can have four, three, two, one, or even no arguments. In the above example, we used one argument (i.e., `before`). The meanings of these arguments are:

- `before`: Whether the chunk hook is executed before or after the code chunk itself is executed. Note that a chunk hook is executed twice for every code chunk (once before with `hook(before = TRUE)` and once after with `hook(before = FALSE`).

- `options`: The list of chunk options for the current code chunk, e.g., `list(fig.width = 5, echo = FALSE, ...)`.

- `envir`: The environment in which the chunk hook is evaluated.

- `name`: The name of the chunk option that triggered the chunk hook.

As we mentioned in the beginning of this chapter, non-character values returned by chunk hooks are silently ignored, and character values are written to the output document.

## Crop plots {#crop-plot}

The chunk hook `knitr::hook_pdfcrop()` can be used to crop PDF and other types of plot files, i.e., remove the extra margins in plots. To enable it, set this hook via `knit_hooks$set()` in a code chunk, and turn on the corresponding chunk option, e.g.,

```{r}
knitr::knit_hooks$set(crop = knitr::hook_pdfcrop)
```

Then you can use the chunk option `crop = TRUE` to crop plots in a code chunk.

The hook `hook_pdfcrop()` calls the external program `pdfcrop` to crop PDF files. This program often comes with a LaTeX distribution (e.g., TeX Live or MiKTeX). You can check if it is available in your system via:

```{r}
# if the returned value is not empty, it is available
Sys.which('pdfcrop')
```

If you are using the LaTeX distribution TinyTeX (see Section \@ref(install-latex)), and `pdfcrop` is not available in your system, you may install it via `tinytex::tlmgr_install('pdfcrop')`.

For non-PDF plot files such as PNG or JPEG files, this hook function calls the R package **magick** to crop them. You need to make sure this R package has been installed. Figure \@ref(fig:crop-no) shows a plot that is not cropped, and Figure \@ref(fig:crop-yes) shows the same plot but has been cropped.

```{r, crop-no, crop=NULL, echo=FALSE, fig.cap='A plot that is not cropped.'}
par(bg = 'gray', fg = 'yellow')
plot(cars)
```

```{r, crop-yes, crop=TRUE, echo=FALSE, fig.cap='A plot that is cropped.', ref.label='crop-no'}
```

## Report how much time each chunk takes to run {#time-chunk}

By default, **knitr** provides a text-based progress bar to show you the knitting progress. If you want more precise timing information about the chunks, you may register a custom chunk hook to record the time for each chunk. Here is an example hook:

```{r, eval=FALSE}
knitr::knit_hooks$set(time_it = local({
  now = NULL
  function(before, options) {
    if (before) {
      # record the current time before each chunk
      now <<- Sys.time()
    } else {
      # calculate the time difference after a chunk
      res <- difftime(Sys.time(), now)
      # return a character string to show the time
      paste('Time for this code chunk to run:', res)
    }
  }})
)
```

Then you can time a chunk with the chunk option `time_it`, e.g.,

````
```{r, time_it = TRUE}`r ''`
Sys.sleep(2)
```
````

If you want to time all code chunks, you can certainly set the option globally: `knitr::opts_chunk$set(time_it = TRUE)`.

In the above hook function, you can also output more information from the chunk options (i.e., the `options` argument of the function). For example, you may print out the chunk label in the returned value:

```{r, eval=FALSE}
paste('Time for the chunk', options$label, 'to run:', res)
```

Or you may record the time without printing it out in the hook:

```{r, eval=FALSE}
all_times <- list()  # store the time for each chunk
knitr::knit_hooks$set(time_it = local({
  now = NULL
  function(before, options) {
    if (before) {
      now <<- Sys.time()
    } else {
      res <- difftime(Sys.time(), now)
      all_times[[options$label]] <<- res
    }
  }})
)
```

Then you can access all the time information in the object `all_times`, which is a named list with the names being chunk labels, and element values being the execution time for each chunk.
