# Workflow

Tips for running your R Markdown projects.

## Creating custom render functions {#customRender}


## Multi-file documents

<!-- https://stackoverflow.com/questions/25824795/how-to-combine-two-rmarkdown-rmd-files-into-a-single-output -->

With larger R Markdown projects, it can be useful to split the document into smaller files which are easier to handle than a single, massive `'.Rmd` document. Also, we may also have generic components which we wish to use across all of our analyses, and rather than copying and pasting the same code across for each report.

In this section, we highlight two approaches which can be used when splitting an R Markdown document into separate `.Rmd` files, using child documents and the **bookdown** [@R-bookdown] package. Most importantly, we will highlight when it is preferred for each to be used within your analysis.

### Child documents

We can include other R Markdown documents within a file by using child documents. The contents of the child document is loaded into the parent document, and any R code will be run as normally within **knitr**.

To include a child document, we can add a code chunk with the option `child`, providing the file path to the `.Rmd` file to be included. The chunk should contain no code but we can specify additional chunk options if we wish to suppress the code output i.e. `echo=FALSE`. For example:

````md
```{r, child = "about.Rmd"}`r ''`
```
````

Whereby the content of `about.Rmd` provides a short description of the packages used for the analysis:

```md
# About

This analysis was written in the R programming language [@R-base]. The report was generated using **rmarkdown** [@R-markdown] and **knitr** [@R-knitr].
```

Child documents are useful where we have a standard piece of text that we wish to include across multiple documents. Rather than copying and pasting the text to each file, we can refer to a single source file. If any changes are made to our generic description at a later date, we can update a single source file and all the reports will be up to date. 

Further examples where child documents could be useful include:

- **Project Description**: it is often useful to include generic description of the project
- **Funding Statements**: research will often have the same funding statement used across all outputs
- **R Session Information**: you may wish to include document metadata

### Using bookdown

The **bookdown** [@R-bookdown] package is designed for creating long format documents within R Markdown, and allows us to split our files into multiple components. To demonstrate the use, we provide a minimal example comprising of three files within the same directory:

```md
directory
  |- index.Rmd
  |- 01-intro.Rmd
  |- 02-analysis.Rmd
```
- **index.Rmd**:

The main detail is that the main file has to be called `index.Rmd` , and must include the additional YAML line site: `bookdown::bookdown_site`. We must also use a **bookdown** output format such as `html_document2`, `pdf_document2`, `word_document2`:

````md
---
title: "A Minimal bookdown document"
site: bookdown::bookdown_site
output: bookdown::pdf_document2
---

Some content
````

- **01-intro.Rmd:**

````md
# Chapter 1

This is chapter 1.
````

- **02-analysis.Rmd**:

```md
# Chapter 2

This is chapter 2.
```

When the file is built, **bookdown** merges the results of the separate files into a single output document. This will automatically be done in alphabetical order, so it makes sense to precede your file name with a numeric index as above.

For a comprehensive overview of the package, readers may refer to the [*Bookdown Book*]() [@xie2016bookdown] or [Chapter 18](https://bookdown.org/yihui/rmarkdown/) of the *R Markdown Definitive Guide* [@xie2018].

### Which method should you use?

There is a legacy of using child documents when creating longer documents, as **bookdown** was only released in 2016. As such, there is some confusion over which should be used when. In general, if you are wishing to reuse a component across multiple documents, it is recommended to use child documents. However, if you are wanting to write a book and split it into multiple files (preferably one per chapter), **bookdown** provides a lot of extra functionality and is generally preferred.

## RStudio integration with R Markdown

R Markdown format and features can be used with R and **rmarkdown** package, with any editor that got your preference. However, RStudio has a deep integration with R Markdown and offers additional feature that will help your project workflow without a doubt.

### R Markdown specific Rstudio keyboard shorcuts

<!-- https://stackoverflow.com/questions/30938280/keyboard-shortcut-to-produce-code-chunk-brackets-in-markdown-in-r-for-rstudio -->

Like any great IDE, RStudio has keyboard shorcuts ^[A full list can be found on the [web](https://support.rstudio.com/hc/en-us/articles/200711853-Keyboard-Shortcuts) or directly in the IDE under tools menu `Tools -> Keyboard Shortcuts Help.`] and some are specific to R Markdown. Some of the most useful shortcuts are summarised in Table \@ref(tab:keyboardShortcuts).

```{r, include = FALSE}
ks_win <- function(letters, ctrl = TRUE, alt = TRUE, shift = FALSE, enter = FALSE) {
  paste0(
    if (ctrl) "Ctrl+",
    if (alt) "Alt+",
    if (shift) "Shift+",
    if (enter) "Enter+",
    letters
  )
}
ks_mac <- function(letters, cmd = TRUE, opt = TRUE, shift = FALSE, enter = FALSE) {
  paste0(
    if (cmd) "Command+",
    if (opt) "Option+",
    if (shift) "Shift+",
    if (enter) "Enter+",
    letters
  )
}
```

```{r keyboardShortcuts, echo = FALSE}
keyboard_table <- tibble::tribble(
  ~ "Description"             , ~ "Windows & Linux"                    , ~ "Mac",
  "Insert R chunk"            , ks_win("I")                            , ks_mac("I"),
  "Preview HTML"              , ks_win("K", alt = FALSE, shift = TRUE) , ks_mac("K", opt = FALSE, shift = TRUE),
  "Knitr document (knitr)"    , ks_win("K", alt = FALSE, shift = TRUE) , ks_mac("K", opt = FALSE, shift = TRUE),
  "Compile Notebook"          , ks_win("K", alt = FALSE, shift = TRUE) , ks_mac("K", opt = FALSE, shift = TRUE),
  "Compile PDF"               , ks_win("K", alt = FALSE, shift = TRUE) , ks_mac("K", opt = FALSE, shift = TRUE),
  "Run all chunks above"      , ks_win("P")                            , ks_mac("P"),
  "Run current chunk"         , ks_win("C")                            , ks_mac("C"),
  "Run current chunk"         , ks_win("Enter", TRUE, FALSE, TRUE)     , ks_mac("Enter", TRUE, FALSE, TRUE),
  "Run next chunk"            , ks_win("N")                            , ks_mac("N"),
  "Run all chunks"            , ks_win("R")                            , ks_mac("R"),
  "Go to next chunk/title"    , ks_win("PgDown", alt = FALSE)          , ks_mac("PgDown", opt = FALSE),
  "Go to previous chunk/title", ks_win("PgUp", alt = FALSE)            , ks_mac("PgUp", opt = FALSE),
  "Show/hide document outline", ks_win("O", TRUE, FALSE, TRUE)         , ks_mac("O", TRUE, FALSE, TRUE),
  "Build book, website, ..."  , ks_win("B", TRUE, FALSE, TRUE)         , ks_mac("B", TRUE, FALSE, TRUE)
)
knitr::kable(keyboard_table, caption = "Rmarkdown related Rstudio keyboard shorcuts")
```

### Customising the knit button

When you click the `Knit` button in RStudio, it will run the `rmarkdown::render` in a clean R session and output a file of the same name in the same directory. For example, running `example.Rmd` with the `output:html_document` will create an output file `example.html`. 

There may be situations where we want to customise how the document is rendered. For example, perhaps we would like the rendered document to contain the current date, or would like to output the compiled report into a different directory. Although we can customise the `render` function using external scripts (See section \@ref(customRender)), it can be inconvenient to have to rely on external scripts to compile your report.

It is possible to control the behaviour of the knit button by using the `knit` argument within the YAML of your document. This requires a function to be specified with the arguments `inputFile`, and `encoding`.  There are two techniques that we can use to customise the behaviour of the knit function:

1. Embed the custom function directly within the YAML
2. Create a custom function and save it in a package

The easiest way of creating a function is to store the code directly within the YAML. We must wrap the entire function in brackets, and any linebreaks in the YAML must start with 4 spaces.  For example, if we want the output filename to include the date it was rendered, we could use the following YAML code:

```yaml
---
knit: (function(inputFile, encoding)
    {rmarkdown::render(input = inputFile,
    encoding = encoding,
    output_file = paste0(tools::file_path_sans_ext(basename(inputFile)),
                                         '-',
                                         format(Sys.Date(), "%Y%m%d"),
                                         '.html'))})
---
``` 

For example, we if knitted `example.Rmd` on `r format(Sys.Date(), '%Y/%m/%d')`, the output file name would be `r paste0("example", '-', format(Sys.Date(), '%Y%m%d'), ".html")`.

While the above approach does benefit from simplicity, it may get more complicated if we wish to do more advanced behaviour (perhaps creating a directory, setting some other environment variables). Embedding longer functions directly in your YAML may become difficult to manage. It may therefore be worth creating a function which could be used in a package. Taking our example above, we have created an example `knit_with_date` as defined below:

```r
#' Custom Knit function for RStudio
#'
#' @export
knit_with_date <- function(inputFile, encoding) {

  rmarkdown::render(inputFile,
                    encoding = encoding,
                    output_file = paste0(dirname(inputFile),'/', tools::file_path_sans_ext(basename(inputFile)),
                                         '-',
                                         format(Sys.Date(), '%Y%m%d'),
                                         '.html'))
  }
```

If we were to add this to the package `myPackage`, you will then be able to refer to your custom Knit function using the following YAML argument:
  
```yaml
---
knit: myPackage::knit_with_date
---
```
                                    
Readers may wish to refer to Section \@ref(customRender) to find out more ideas on how they could customise the control of their render function.

## Writing to file from an Rmd file with the `cat` engine

<!-- https://github.com/dr-harper/rmarkdown-cookbook/issues/140 -->

Sometimes it could be useful to write to some files from inside the Rmd. If this
could be achieved using pure R code in a r code chunk, `knitr` as a little
unknown gem for this : the `cat` engine.

### About the `cat` engine 

The `cat` engine allows to write to a file some codes from a Rmarkdown chunk.
The file path is passed as an argument in the generic `engine.opts` chunk
options: `engine.opts = list(file = <file to write to>)`. The element in this
list will be pass to `base::cat()` function allowing to customize how the
content is written to file.

In this example, we will write to a temp file with `fs::path_temp()`. 

````md
```{cat, engine.opts = list(file = fs::path_temp('script.R'))}`r ''`
my_function <- function(x){ x+1 }
```
```` 

```{cat, engine.opts = list(file = fs::path_temp('script.R'))}
my_function <- function(x){ x+1 }
```

The content can be used by filling a chunk from that file. This is another trick made possible using the `code` chunk option. You'll see the content of the file

````md
```{r, code = readLines(fs::path_temp('script.R'))}`r ''`
```
````

This will render

```{r, code = readLines(fs::path_temp('script.R'))}
```

As the code above is executed, the function is available in any following r chunk.

```{r, results='hold'}
my_function(5)
rm(my_function)
```

And the file can be sourced as usual too.

```{r, results='hold'}
source(fs::path_temp('script.R'))
my_function(5)
rm(my_function)
```


Writing to an R script is not really a real use case example and the `cat`
engine can handled any format.

Here, we use it to write some yaml to a file. The default behavior for `cat`
engine is to show nothing in the output document. However, here we will also
show the content in the output R Markdown file. The `cat` engine understand the
field `lang` in `engine.opts` option or directly `class.source` chunk option for
html outputs (see Section \@ref(chunk-styling)). You can provide any language
supported by the syntax highlighting. This two chunks are equivalent:

````md
```{cat, engine.opts = list(file = fs::path_temp('demo.yml'), lang = "yaml")}`r ''`
a:
  aa: something
  bb: 1
b:
  aa: somethingelse
  bb: 2
```
````

````md
```{cat, engine.opts = list(file = fs::path_temp('demo.yml')), class.source = "yaml"}`r ''`
a:
  aa: something
  bb: 1
b:
  aa: somethingelse
  bb: 2
```
````

They will write to file and print the code with the chosen syntax highlighting, here `yaml`
```{cat, engine.opts = list(file = fs::path_temp('demo.yml'), lang = "yaml")}
a:
  aa: something
  bb: 1
b:
  aa: somethingelse
  bb: 2
```

As previously, the file can be access later on and read into R
```{r}
yaml::read_yaml(fs::path_temp('demo.yml'))
```


```{r cleaning, include = FALSE}
unlink(c(fs::path_temp('demo.yml'), fs::path_temp('script.R')))
```

### Generate a template file from inside the rmarkdown

<!-- https://github.com/dr-harper/rmarkdown-cookbook/issues/149 -->

As detailed in section \@ref(chunk-styling), `css` chunk can directly be used to
apply custom style. However, it is also possible to provide a custom css file to
pandoc directly to be used with the html output and the `cat` engine can be used
to write this css file from the Rmarkdown.

One important key is that all chunks are executed before pandoc conversion. This
means it is possible in the yaml header to link to a file that does not exist
and will be created when rendering the document.

This complete example shows how to generate `my_custom.css` from a chunk in the
document while telling pandoc to use this css with the yaml field `css:
my_custom.css`

```{r, echo = FALSE, comment=""}
import_example("generate-css-cat-engine.Rmd")
```

This could also be used to generate any file that can be given to pandoc, like
an `in_header` tex file for a `pdf_document` output.^[Those yaml options for
pandoc can be found in *R Markdown Definitive Guide* [@xie2018]]

```{r, echo = FALSE, comment=""}
import_example("generate-latex-header-with-cat-engine.Rmd")
```

