# Knitr

As illustrated in Figure \@ref(fig:rmdworkflow), the R package **knitr** is one of the two major components of R Markdown (the other being Pandoc). In this chapter, we show some recipes related to **knitr**.

## Using knitr options

There are more than 40 different chunk options which can be used to fine-tune the behaviour of **knitr** when processing R chunks. [Section 2.8](https://bookdown.org/yihui/rmarkdown/r-code.html) of @xie2018 provides a subset of these, while the full list is available at <https://yihui.name/knitr/options/>.

This section aims to highlight some useful **knitr** chunk options that may not be widely known. In the following recipes, we only show examples of applying chunk options to individual code chunks. However, please be aware of the fact that any chunk options can also be applied globally to a whole document, so you do not have to repeat the options in every single code chunk. To set chunk options globally, call `knitr::opts_chunk$set()` in a code chunk (usually the first one in the document), e.g.,

````md
`r ''````{r, include=FALSE}
knitr::opts_chunk$set(comment = "", echo = FALSE)
```
````

### Hide code and all outputs

We often wish to setup our R Markdown documents by loading a series of packages. Instead of using separate commands to R source code (`echo=FALSE`), hide text output (`results='hide'`), hide warnings (`warnings=FALSE`), hide messages (`messages=FALSE`), and so on, we can suppress the output of the entire code chunk using the single option `include=FALSE` as shown below:

````md
`r ''````{r, include=FALSE}
# any R code here
```
````

With `include=FALSE`, the code chunk will be evaluated (unless `eval=FALSE`), but the output will be completely suppressed.

### Remove leading hashes in text output

<!-- https://stackoverflow.com/questions/15081212/remove-hashes-in-r-output-from-r-markdown-and-knitr -->

By default, R code output will have two hashes `#` inserted in front of the text output. We can alter this behaviour by using the `comment` chunk option, which defaults to a character string of two hashes `"##"`. We can use an empty string if we want to remove the hashes. For example:

````md
`r ''````{r, comment=""}
1:100
```
````

## Exit knitting early

Sometimes we may want to exit knitting early and not at the end of the document. For example, we may be working on some analysis and only wish to share the first half of the results, or we may still be working on code at the bottom that is not yet complete. In these situations, we could consider using the `knit_exit()` function, which will end the knitting process at that location.

Below is a simple example, where we have a very simple chunk followed by a more time-consuming one:

````md
`r ''````{r}
1 + 1
knitr::knit_exit()
```

Okay, you will not see me in the output.

`r ''````{r}
Sys.sleep(100)
```
````

Normally you have to wait for 100 seconds, but since we have called `knit_exit()`, the rest of the document will be ignored.

## Generate a plot and display it elsewhere

Normally plots generated in a code chunk are displayed beneath the code chunk, but you can choose to show them elsewhere and (optionally) hide them in the code chunk. Below is an example:

````md
We generate a plot in this code chunk but do not show it:

`r ''````{r cars-plot, dev='png', fig.show='hide'}
plot(cars)
```

After another paragraph, we introduce the plot:

![A nice plot.](`r knitr::inline_expr("knitr::fig_chunk('cars-plot', 'png')")`)
````

In the code chunk, we used the chunk option `fig.show='hide'` to hide the plot temporarily. Then in another paragraph, we called the function `knitr::fig_chunk()` to retrieve the path of the plot file, which is usually like `test_files/figure-html/cars-plot-1.png`. You need to pass the chunk label and the graphical device name to `fig_chunk()` for it to calculate the plot file path. You may see https://stackoverflow.com/a/46305297/559676 for an application of `fig_chunk()` to **blogdown** websites.

## Allow duplicate labels in code chunks

<!-- https://stackoverflow.com/questions/36868287/purl-within-knit-duplicate-label-error/47065392#47065392 -->

By default, **knitr** does not allow duplicate code chunk labels in the document. Duplicate labels will result in an error when the document is knitted. This occurs most frequently when a code chunk is copied and pasted within a document. You may have seen an error message like this:

```text
processing file: myfile.Rmd
Error in parse_block(g[-1], g[1], params.src) :
  duplicate label 'cars'
Calls: <Anonymous> ... process_file -> split_file -> lapply ->
  FUN -> parse_block
Execution halted
```

However, there are scenarios where we may wish to allow duplicate labels. For example, if we have one master document `parent.Rmd` in which we knit the child document multiple times, it will fail:

```{r, eval = FALSE}
# settings
settings = list(...)

# run once
knit_child('useful_analysis.rmd')

# new settings
settings = list(...)

# run again
knit_child('useful_analysis.rmd')
```

In this scenario, we can allow duplicate labels by setting this global option in R before the child document is knitted:

```{r, eval = FALSE}
options(knitr.duplicate.label = 'allow')
```

You should set this option with caution. As with most error messages, they are there for a reason. Allowing duplicate chunks can create silent problems with figures and cross-references. For example, if two code chunks have the same label and both chunks generate plots, their plot files will overwrite each other (without an error or warning message), because the filenames of plots are determined by the chunk labels.
