# Other Languages {#other-languages}


## Write the chunk content to a file via the `cat` engine {#eng-cat}

<!-- https://github.com/yihui/rmarkdown-cookbook/issues/140 -->

Sometimes it could be useful to write to some files from inside the Rmd. If this
could be achieved using pure R code in a r code chunk, `knitr` as a little
unknown gem for this : the `cat` engine.

### About the `cat` engine 

The `cat` engine allows to write to a file some codes from a Rmarkdown chunk.
The file path is passed as an argument in the generic `engine.opts` chunk
options: `engine.opts = list(file = <file to write to>)`. The element in this
list will be pass to `base::cat()` function allowing to customize how the
content is written to file.

In this example, we will write to a temp file with `fs::path_temp()`. 

````md
```{cat, engine.opts = list(file = fs::path_temp('script.R'))}`r ''`
my_function <- function(x){ x+1 }
```
```` 

```{cat, engine.opts = list(file = fs::path_temp('script.R'))}
my_function <- function(x){ x+1 }
```

The content can be used by filling a chunk from that file. This is another trick made possible using the `code` chunk option. You'll see the content of the file

````md
```{r, code = readLines(fs::path_temp('script.R'))}`r ''`
```
````

This will render

```{r, code = readLines(fs::path_temp('script.R'))}
```

As the code above is executed, the function is available in any following r chunk.

```{r, results='hold'}
my_function(5)
rm(my_function)
```

And the file can be sourced as usual too.

```{r, results='hold'}
source(fs::path_temp('script.R'))
my_function(5)
rm(my_function)
```


Writing to an R script is not really a real use case example and the `cat`
engine can handle any format.

Here, we use it to write some yaml to a file. The default behavior for `cat`
engine is to show nothing in the output document. However, here we will also
show the content in the output R Markdown file. The `cat` engine understand the
field `lang` in `engine.opts` option or directly `class.source` chunk option for
html outputs (see Section \@ref(chunk-styling)). You can provide any language
supported by the syntax highlighting. These two chunks are equivalent:

````md
```{cat, engine.opts = list(file = fs::path_temp('demo.yml'), lang = "yaml")}`r ''`
a:
  aa: something
  bb: 1
b:
  aa: somethingelse
  bb: 2
```
````

````md
```{cat, engine.opts = list(file = fs::path_temp('demo.yml')), class.source = "yaml"}`r ''`
a:
  aa: something
  bb: 1
b:
  aa: somethingelse
  bb: 2
```
````

They will write to file and print the code with the chosen syntax highlighting, here `yaml`
```{cat, engine.opts = list(file = fs::path_temp('demo.yml'), lang = "yaml")}
a:
  aa: something
  bb: 1
b:
  aa: somethingelse
  bb: 2
```

As previously, the file can be accessed later on and read into R
```{r}
yaml::read_yaml(fs::path_temp('demo.yml'))
```


```{r cleaning, include = FALSE}
unlink(c(fs::path_temp('demo.yml'), fs::path_temp('script.R')))
```

### Generate a template file from inside the rmarkdown

<!-- https://github.com/yihui/rmarkdown-cookbook/issues/149 -->

As detailed in section \@ref(chunk-styling), `css` chunk can directly be used to
apply custom style. However, it is also possible to provide a custom css file to
pandoc directly to be used with the html output and the `cat` engine can be used
to write this css file from the Rmarkdown.

One important key is that all chunks are executed before pandoc conversion. This
means it is possible in the yaml header to link to a file that does not exist
and will be created when rendering the document.

This complete example shows how to generate `my_custom.css` from a chunk in the
document while telling pandoc to use this css with the yaml field `css:
my_custom.css`

```{r, echo = FALSE, comment=""}
import_example("generate-css-cat-engine.Rmd")
```

This could also be used to generate any file that can be given to pandoc, like
an `in_header` tex file for a `pdf_document` output.^[Those yaml options for
pandoc can be found in *R Markdown Definitive Guide* [@rmarkdown2018]]

```{r, echo = FALSE, comment=""}
import_example("generate-latex-header-with-cat-engine.Rmd")
```

