# Workflow

In this chapter, we introduce some tips on working with individual R Markdown documents as well as running your R Markdown projects.

## RStudio keyboard shortcuts {#rstudio-shortcuts}

The R Markdown format can be used with any editor of your choice, as long as R, the **rmarkdown** package, and Pandoc are installed. However, RStudio has a deep integration with R Markdown so you can work with R Markdown smoothly.

Like any IDE, RStudio has keyboard shortcuts. A full list can be found under the menu `Tools -> Keyboard Shortcuts Help.` Some of the most useful shortcuts related to R Markdown are summarized in Table \@ref(tab:shortcuts).

```{r, include = FALSE}
ks_win <- function(letters, ctrl = TRUE, alt = TRUE, shift = FALSE, enter = FALSE) {
  paste0(
    if (ctrl) "Ctrl+",
    if (alt) "Alt+",
    if (shift) "Shift+",
    if (enter) "Enter+",
    letters
  )
}
ks_mac <- function(letters, cmd = TRUE, opt = TRUE, shift = FALSE, enter = FALSE) {
  paste0(
    if (cmd) "Command+",
    if (opt) "Option+",
    if (shift) "Shift+",
    if (enter) "Enter+",
    letters
  )
}
```

```{r shortcuts, echo = FALSE}
keyboard_table <- tibble::tribble(
  ~ "Task"             , ~ "Windows & Linux"                    , ~ "macOS",
  "Insert R chunk"            , ks_win("I")                            , ks_mac("I"),
  "Preview HTML"              , ks_win("K", alt = FALSE, shift = TRUE) , ks_mac("K", opt = FALSE, shift = TRUE),
  "Knitr document (knitr)"    , ks_win("K", alt = FALSE, shift = TRUE) , ks_mac("K", opt = FALSE, shift = TRUE),
  "Compile Notebook"          , ks_win("K", alt = FALSE, shift = TRUE) , ks_mac("K", opt = FALSE, shift = TRUE),
  "Compile PDF"               , ks_win("K", alt = FALSE, shift = TRUE) , ks_mac("K", opt = FALSE, shift = TRUE),
  "Run all chunks above"      , ks_win("P")                            , ks_mac("P"),
  "Run current chunk"         , ks_win("C")                            , ks_mac("C"),
  "Run current chunk"         , ks_win("Enter", TRUE, FALSE, TRUE)     , ks_mac("Enter", TRUE, FALSE, TRUE),
  "Run next chunk"            , ks_win("N")                            , ks_mac("N"),
  "Run all chunks"            , ks_win("R")                            , ks_mac("R"),
  "Go to next chunk/title"    , ks_win("PgDown", alt = FALSE)          , ks_mac("PgDown", opt = FALSE),
  "Go to previous chunk/title", ks_win("PgUp", alt = FALSE)            , ks_mac("PgUp", opt = FALSE),
  "Show/hide document outline", ks_win("O", TRUE, FALSE, TRUE)         , ks_mac("O", TRUE, FALSE, TRUE),
  "Build book, website, ..."  , ks_win("B", TRUE, FALSE, TRUE)         , ks_mac("B", TRUE, FALSE, TRUE)
)
knitr::kable(keyboard_table, caption = "RStudio keyboard shortcuts related to R Markdown.", booktabs = TRUE)
```

Besides, you can press `F7` to spellcheck your document. You can also restart the R session by `Ctrl + Alt + F10` (or `Command + Option + F10` on macOS), which can be helpful for reproducibility, because results are more likely to be reproducible if they are computed from a new R session. This can also be done through the dropdown menu "Restart R and Run All Chunks" behind the "Run" button on the toolbar.

## Customizing the `Knit` button {#custom-knit}

When you click the `Knit` button in RStudio, it will run the `rmarkdown::render` in a clean R session and output a file of the same name in the same directory. For example, running `example.Rmd` with the `output:html_document` will create an output file `example.html`. 

There may be situations where we want to customize how the document is rendered. For example, perhaps we would like the rendered document to contain the current date, or would like to output the compiled report into a different directory. Although we can customize the `render` function using external scripts (See section \@ref(customRender)), it can be inconvenient to have to rely on external scripts to compile your report.

It is possible to control the behavior of the knit button by using the `knit` argument within the YAML of your document. This requires a function to be specified with the arguments `inputFile`, and `encoding`.  There are two techniques that we can use to customize the behavior of the knit function:

1. Embed the custom function directly within the YAML
2. Create a custom function and save it in a package

The easiest way of creating a function is to store the code directly within the YAML. We must wrap the entire function in brackets, and any linebreaks in the YAML must start with 4 spaces.  For example, if we want the output filename to include the date it was rendered, we could use the following YAML code:

```yaml
---
knit: (function(inputFile, encoding)
    {rmarkdown::render(input = inputFile,
    encoding = encoding,
    output_file = paste0(tools::file_path_sans_ext(basename(inputFile)),
                                         '-',
                                         format(Sys.Date(), "%Y%m%d"),
                                         '.html'))})
---
``` 

For example, we if knitted `example.Rmd` on `r format(Sys.Date(), '%Y/%m/%d')`, the output file name would be `r paste0("example", '-', format(Sys.Date(), '%Y%m%d'), ".html")`.

While the above approach does benefit from simplicity, it may get more complicated if we wish to do more advanced behavior (perhaps creating a directory, setting some other environment variables). Embedding longer functions directly in your YAML may become difficult to manage. It may therefore be worth creating a function which could be used in a package. Taking our example above, we have created an example `knit_with_date` as defined below:

```r
#' Custom Knit function for RStudio
#'
#' @export
knit_with_date <- function(inputFile, encoding) {

  rmarkdown::render(inputFile,
                    encoding = encoding,
                    output_file = paste0(dirname(inputFile),'/', tools::file_path_sans_ext(basename(inputFile)),
                                         '-',
                                         format(Sys.Date(), '%Y%m%d'),
                                         '.html'))
  }
```

If we were to add this to the package `myPackage`, you will then be able to refer to your custom Knit function using the following YAML argument:
  
```yaml
---
knit: myPackage::knit_with_date
---
```
                                    
Readers may wish to refer to Section \@ref(customRender) to find out more ideas on how they could customize the control of their render function.

## Writing to file from an Rmd file with the `cat` engine

<!-- https://github.com/dr-harper/rmarkdown-cookbook/issues/140 -->

Sometimes it could be useful to write to some files from inside the Rmd. If this
could be achieved using pure R code in a r code chunk, `knitr` as a little
unknown gem for this : the `cat` engine.

### About the `cat` engine 

The `cat` engine allows to write to a file some codes from a Rmarkdown chunk.
The file path is passed as an argument in the generic `engine.opts` chunk
options: `engine.opts = list(file = <file to write to>)`. The element in this
list will be pass to `base::cat()` function allowing to customize how the
content is written to file.

In this example, we will write to a temp file with `fs::path_temp()`. 

````md
```{cat, engine.opts = list(file = fs::path_temp('script.R'))}`r ''`
my_function <- function(x){ x+1 }
```
```` 

```{cat, engine.opts = list(file = fs::path_temp('script.R'))}
my_function <- function(x){ x+1 }
```

The content can be used by filling a chunk from that file. This is another trick made possible using the `code` chunk option. You'll see the content of the file

````md
```{r, code = readLines(fs::path_temp('script.R'))}`r ''`
```
````

This will render

```{r, code = readLines(fs::path_temp('script.R'))}
```

As the code above is executed, the function is available in any following r chunk.

```{r, results='hold'}
my_function(5)
rm(my_function)
```

And the file can be sourced as usual too.

```{r, results='hold'}
source(fs::path_temp('script.R'))
my_function(5)
rm(my_function)
```


Writing to an R script is not really a real use case example and the `cat`
engine can handle any format.

Here, we use it to write some yaml to a file. The default behavior for `cat`
engine is to show nothing in the output document. However, here we will also
show the content in the output R Markdown file. The `cat` engine understand the
field `lang` in `engine.opts` option or directly `class.source` chunk option for
html outputs (see Section \@ref(chunk-styling)). You can provide any language
supported by the syntax highlighting. These two chunks are equivalent:

````md
```{cat, engine.opts = list(file = fs::path_temp('demo.yml'), lang = "yaml")}`r ''`
a:
  aa: something
  bb: 1
b:
  aa: somethingelse
  bb: 2
```
````

````md
```{cat, engine.opts = list(file = fs::path_temp('demo.yml')), class.source = "yaml"}`r ''`
a:
  aa: something
  bb: 1
b:
  aa: somethingelse
  bb: 2
```
````

They will write to file and print the code with the chosen syntax highlighting, here `yaml`
```{cat, engine.opts = list(file = fs::path_temp('demo.yml'), lang = "yaml")}
a:
  aa: something
  bb: 1
b:
  aa: somethingelse
  bb: 2
```

As previously, the file can be accessed later on and read into R
```{r}
yaml::read_yaml(fs::path_temp('demo.yml'))
```


```{r cleaning, include = FALSE}
unlink(c(fs::path_temp('demo.yml'), fs::path_temp('script.R')))
```

### Generate a template file from inside the rmarkdown

<!-- https://github.com/dr-harper/rmarkdown-cookbook/issues/149 -->

As detailed in section \@ref(chunk-styling), `css` chunk can directly be used to
apply custom style. However, it is also possible to provide a custom css file to
pandoc directly to be used with the html output and the `cat` engine can be used
to write this css file from the Rmarkdown.

One important key is that all chunks are executed before pandoc conversion. This
means it is possible in the yaml header to link to a file that does not exist
and will be created when rendering the document.

This complete example shows how to generate `my_custom.css` from a chunk in the
document while telling pandoc to use this css with the yaml field `css:
my_custom.css`

```{r, echo = FALSE, comment=""}
import_example("generate-css-cat-engine.Rmd")
```

This could also be used to generate any file that can be given to pandoc, like
an `in_header` tex file for a `pdf_document` output.^[Those yaml options for
pandoc can be found in *R Markdown Definitive Guide* [@rmarkdown2018]]

```{r, echo = FALSE, comment=""}
import_example("generate-latex-header-with-cat-engine.Rmd")
```

