# (APPENDIX) Appendix {-}

# Chunk and Package Options {#chunk-options-all}

The **knitr** package has provided a lot of chunk options for customizing nearly
all components of code chunks, such as the source code, text output, plots, and
the language of the chunk, etc. It has also offered some options at the package
level to customize the knitting process. This page documents all chunk options
and package options available in **knitr**. The default values of these options
are in parentheses in the list items.

Please see Chapter \@ref(chunk-options) for specific examples and recipes using these options.

An up-to-date version of this document is maintained online at https://yihui.org/knitr/options/. 

## Chunk Options {#chunk-options-details}

Chunk options are written in chunk headers. The syntax for chunk headers depends
on the document format, e.g., for `.Rnw` documents (R + LaTeX), chunk headers
are written with `<< >>=`, and for `.Rmd` documents, chunk headers are written
with ```` ```{r} ````. The examples on this page are primarily for `.Rmd`
documents (R Markdown), but in most cases, the chunk options can be used with
any document formats.

Chunk options are written in the form `tag=value` like this:

```` md
```{r, my-chunk, echo=FALSE, fig.height=4, dev='jpeg'}`r ''`
```
````

A special chunk option is the chunk label (e.g., `my-chunk` in the above
example). Only the chunk label does not need a `tag` (i.e., you only provide the
`value`). If you prefer the form `tag=value`, you could also use the chunk
option `label` explicitly, e.g.,

```` md
```{r, label='my-chunk'}`r ''`
```
````

The chunk label for each chunk is assumed to be unique within the document. This
is especially important for cache and plot filenames, because these filenames
are based on chunk labels. Chunks without labels will be assigned labels like
`unnamed-chunk-i`, where `i` is an incremental number.

You may use `knitr::opts_chunk$set()` to change the default values of chunk
options in a document. For example, you may put this in the first code chunk of
your document:

```` md
```{r, setup, include=FALSE}`r ''`
knitr::opts_chunk$set(
  comment = '', fig.width = 6, fig.height = 6
)
```
````

Below are a few more tips about chunk options:

1.  The chunk header must be written on one line. You must not break the line.

2.  Try to avoid spaces, periods (`.`), and underscores (`_`) in chunk labels
    and paths. If you need separators, you are recommended to use hyphens (`-`)
    instead. For example, `setup-options` is a good label, whereas
    `setup.options` and `chunk 1` are bad; `fig.path = 'figures/mcmc-'` is a
    good path for figure output, and `fig.path = 'markov chain/monte carlo'` is
    bad.

3.  All option values must be *valid R expressions*. You may think of them as
    values to be passed to function arguments.

    -   For example, options that take *character* values must be quoted, e.g.,
        `results = 'asis'` and `out.width='\\textwidth'` (remember that a
        literal backslash needs double backslashes).
    -   In theory, the chunk label should be quoted, too. However, for the sake
        of convenience, it will be automatically quoted if you did not (e.g.,
        ```` ```{r, 2a} ```` will parsed as ```` ```{r, '2a'} ````).
    -   You can write arbitrarily complicated expressions as long as they are
        valid R code.

Below is a list of chunk options in **knitr** documented in the format
"`option`: (`default value`; type of value)".

### Code evaluation

-   `eval`: (`TRUE`; logical or numeric) Whether to evaluate the code chunk. It
    can also be a numeric vector to choose which R expression(s) to evaluate,
    e.g., `eval = c(1, 3, 4)` will evaluate the first, third, and forth
    expressions, and `eval = -(4:5)` will evaluate all expressions except the
    forth and fifth.

### Text output

-   `echo`: (`TRUE`; logical or numeric) Whether to display the source code in
    the output document. Besides `TRUE`/`FALSE`, which shows/hides the source
    code, we can also use a numeric vector to choose which R expression(s) to
    echo in a chunk, e.g., `echo = 2:3` means to echo only the 2nd and 3rd
    expressions, and `echo = -4` means to exclude the 4th expression.

-   `results`: (`'markup'`; character) Controls how to display the text results.
    Note that this option only applies to normal text output (not warnings,
    messages, or errors). The possible values are:

    -   `markup`: Mark up text output with the appropriate environments
        depending on the output format. For example, for R Markdown, if the text
        output is a character string `"[1] 1 2 3"`, the actual output that
        **knitr** produces will be:

        ```` md
        ```
        [1] 1 2 3
        ```
        ````

        In this case, `results='markup'` means to put the text output in fenced
        code blocks (```` ``` ````).

    -   `asis`: Write text output as-is, i.e., write the raw text results
        directly into the output document without any markups.

        ```` md
        ```{r, results='asis'}
        cat("I'm raw **Markdown** content.\n")
        ```
        ````

    -   `hold`: Hold all pieces of text output in a chunk and flush them to the
        end of the chunk.

    -   `hide` (or `FALSE`): Hide text output.

-   `collapse`: (`FALSE`; logical) Whether to, if possible, collapse all the
    source and output blocks from one code chunk into a single block (by
    default, they are written to separate blocks). This option only applies to
    Markdown documents.

-   `warning`: (`TRUE`; logical) Whether to preserve warnings (produced by
    `warning()`) in the output. If `FALSE`, all warnings will be printed in the
    console instead of the output document. It can also take numeric values as
    indices to select a subset of warnings to include in the output. Note that
    these values reference the indices of the warnings themselves (e.g., `3`
    means "the third warning thrown from this chunk") and not the indices of
    which expressions are allowed to emit warnings.

-   `error`: (`TRUE`; logical) Whether to preserve errors (from `stop()`). By
    default, the code evaluation will not stop even in case of errors! If we
    want to stop on errors, we need to set this option to `FALSE`. Note that R
    Markdown has changed this default value to `FALSE`. When the chunk option
    `include = FALSE`, **knitr** will stop on error, because it is easy to
    overlook potential errors in this case.

-   `message`: (`TRUE`; logical) Whether to preserve messages emitted by
    `message()` (similar to the option `warning`).

-   `include`: (`TRUE`; logical) Whether to include the chunk output in the
    output document. If `FALSE`, nothing will be written into the output
    document, but the code is still evaluated and plot files are generated if
    there are any plots in the chunk, so you can manually insert figures later.

-   `strip.white`: (`TRUE`; logical) Whether to remove blank lines in the
    beginning or end of a source code block in the output.

-   `class.output`: (`NULL`; character) A vector of class names to be added to
    the text output blocks. This option only works for HTML output formats in R
    Markdown. For example, with `class.output = c('foo', 'bar')`, the text
    output will be placed in `<pre class="foo bar"></pre>`.

-   `class.message`/`class.warning`/`class.error`: (`NULL`; character) Similar
    to `class.output`, but applied to messages, warnings, and errors in R
    Markdown output. Please see the "Code Decoration" section for `class.source`
    which applies similarly to source code blocks.

-   `attr.output`/`attr.message`/`attr.warning`/`attr.error`: (`NULL`;
    character) Similar to the `class.*` options above, but for specifying
    arbitrary fenced code block attributes for Pandoc; `class.*` is a special
    application of `attr.*`, e.g., `class.source = 'numberLines'` is equivalent
    to `attr.source = '.numberLines'`, but `attr.source` can take arbitrary
    attribute values, e.g., `attr.source = c('.numberLines', 'startFrom="11"')`.

-   `render`: (`knitr::knit_print`; `function(x, options, ...)`) A function to
    print the visible values in a chunk. The value passed to the first argument
    of this function (i.e., `x`) is the value evaluated from each expression in
    the chunk. The list of current chunk options is passed to the argument
    `options`. This function is expected to return a character string. For more
    information, check out the package vignette about custom chunk rendering:
    `vignette('knit_print', package = 'knitr')`.

-   `split`: (`FALSE`; logical) Whether to split the output into separate files
    and include them into LaTeX by `\input{}` or HTML by `<iframe></iframe>`.
    This option only works for `.Rnw`, `.Rtex`, and `.Rhtml` documents.

### Code decoration

-   `tidy`: (`FALSE`) Whether to reformat the R code. Other possible values are:

    -   `TRUE` (equivalent to `tidy = 'formatR'`): use `formatR::tidy_source()`
        to reformat the code;
    -   `'styler'`: use `styler::style_text()` to reformat the code;
    -   A custom function of the form `function(code, ...) {}` to return the
        reformatted code;
    -   If reformatting fails, the original R code will not be changed (with a
        warning).

-   `tidy.opts`: (`NULL`; list) A list of options to be passed to the function
    determined by the `tidy` option, e.g.,
    `tidy.opts = list(blank = FALSE, width.cutoff = 60)` for `tidy = 'formatR'`
    to remove blank lines and try to cut the code lines at 60 characters.

-   `prompt`: (`FALSE`; logical) Whether to add the prompt characters in the R
    code. See `prompt` and `continue` on the help page `?base::options`. Note
    that adding prompts can make it difficult for readers to copy R code from
    the output, so `prompt = FALSE` may be a better choice. This option may not
    work well when the chunk option `engine` is not `R`
    ([\#1274](https://github.com/yihui/knitr/issues/1274)).

-   `comment`: (`'##'`; character) The prefix to be added before each line of
    the text output. By default, the text output is commented out by `##`, so if
    readers want to copy and run the source code from the output document, they
    can select and copy everything from the chunk, since the text output is
    masked in comments (and will be ignored when running the copied text). Set
    `comment = ''` remove the default `##`.

-   `highlight`: (`TRUE`; logical) Whether to syntax highlight the source code.

-   `class.source`: (NULL; character) Class names for source code blocks in the
    output document. Similar to the `class.*` options for output such as
    `class.output`.

-   `attr.source`: (`NULL`; character) Attributes for source code blocks.
    Similar to the `attr.*` options for output such as `attr.output`.

-   `size`: (`'normalsize'`; character) Font size of the chunk output from
    `.Rnw` documents. See [this
    page](https://www.overleaf.com/learn/latex/Font_sizes,_families,_and_styles)
    for possible sizes.

-   `background`: (`'#F7F7F7'`; character) Background color of the chunk output
    of `.Rnw` documents.

-   `indent`: (character) A string to be added to each line of the chunk output.
    Typically it consists of white spaces. This option is assumed to be
    read-only, and **knitr** sets its value while parsing the document. For
    example, for the chunk below, `indent` is a character string of two spaces:

    ```` markdown
      ```{r}
      rnorm(10)
      ```
    ````

### Cache

-   `cache`: (`FALSE`; logical) Whether to cache a code chunk. When evaluating
    code chunks for the second time, the cached chunks are skipped (unless they
    have been modified), but the objects created in these chunks are loaded from
    previously saved databases (`.rdb` and `.rdx` files), and these files are
    saved when a chunk is evaluated for the first time, or when cached files are
    not found (e.g., you may have removed them by hand). Note that the filename
    consists of the chunk label with an MD5 digest of the R code and chunk
    options of the code chunk, which means any changes in the chunk will produce
    a different MD5 digest, hence invalidate the cache. See more information on
    [this page](../demo/cache/).

-   `cache.path`: (`'cache/'`; character) A prefix to be used to generate the
    paths of cache files. For R Markdown, the default value is based on the
    input filename, e.g., the cache paths for the chunk with the label `FOO` in
    the file `INPUT.Rmd` will be of the form `INPUT_cache/FOO_*.*`.

-   `cache.vars`: (`NULL`; character) A vector of variable names to be saved in
    the cache database. By default, all variables created in the current chunks
    are identified and saved, but you may want to manually specify the variables
    to be saved, because the automatic detection of variables may not be robust,
    or you may want to save only a subset of variables.

-   `cache.globals`: (`NULL`; character) A vector of the names of variables that
    are not created from the current chunk. This option is mainly for
    `autodep = TRUE` to work more precisely---a chunk `B` depends on chunk `A`
    when any of `B`'s global variables are `A`'s local variables. In case the
    automatic detection of global variables in a chunk fails, you may manually
    specify the names of global variables via this option (see
    [\#1403](https://github.com/yihui/knitr/issues/1403) for an example).

-   `cache.lazy`: (`TRUE`; logical) Whether to `lazyLoad()` or directly `load()`
    objects. For very large objects, lazyloading may not work, so
    `cache.lazy = FALSE` may be desirable (see
    [\#572](https://github.com/yihui/knitr/issues/572)).

-   `cache.comments`: (`NULL`; logical) If `FALSE`, changing comments in R code
    chunks will not invalidate the cache database.

-   `cache.rebuild`: (`FALSE`; logical) If `TRUE`, reevaluate the chunk even if
    the cache does not need to be invalidated. This can be useful when you want
    to conditionally invalidate the cache, e.g.,
    `cache.rebuild = !file.exists("some-file")` can rebuild the chunk when
    `some-file` does not exist (see
    [\#238](https://github.com/yihui/knitr/issues/238)).

-   `dependson`: (`NULL`; character or numeric) A character vector of chunk
    labels to specify which other chunks this chunk depends on. This option
    applies to cached chunks only---sometimes the objects in a cached chunk may
    depend on other cached chunks, so when other chunks are changed, this chunk
    must be updated accordingly.

    -   If `dependson` is a numeric vector, it means the indices of chunk
        labels, e.g. `dependson = 1` means this chunk depends on the first chunk
        in the document, and `dependson = c(-1, -2)` means it depends on the
        previous two chunks (negative indices stand for numbers of chunks before
        this chunk, and note they are always relative to the current chunk).
    -   Please note this option does not work when set as a global chunk option
        via `opts_chunk$set()`; it must be set as a local chunk option.

-   `autodep`: (`FALSE`; logical) Whether to analyze dependencies among chunks
    automatically by detecting global variables in the code (may not be
    reliable), so `dependson` does not need to be set explicitly.

### Plots

-   `fig.path`: (`'figure/'`; character) A prefix to be used to generate figure
    file paths. `fig.path` and chunk labels are concatenated to generate the
    full paths. It may contain a directory like `figure/prefix-`, the directory
    will be created if it does not exist.

-   `fig.keep`: (`'high'`; character) How plots in chunks should be kept.
    Possible values are:

    -   `high`: Only keep high-level plots (merge low-level changes into
        high-level plots);
    -   `none`: Discard all plots;
    -   `all`: Keep all plots (low-level plot changes may produce new plots);
    -   `first`: Only keep the first plot;
    -   `last`: only keep the last plot;
    -   If set to a numeric vector, the values are indices of (low-level) plots
        to keep.

    Low-level plotting commands include `lines()` and `points()`, etc. To better
    understand `fig.keep`, consider the following chunk:

    ```` md
    ```{r, test-plot}
    plot(1)         # high-level plot
    abline(0, 1)    # low-level change
    plot(rnorm(10)) # high-level plot
    # many low-level changes in a loop (a single R expression)
    for(i in 1:10) {
        abline(v = i, lty = 2)
    }
    ```
    ````

    Normally this produces 2 plots in the output (because `fig.keep = 'high'`).
    For `fig.keep = 'none'`, no plots will be saved. For `fig.keep = 'all'`, 4
    plots are saved. For `fig.keep = 'first'`, the plot produced by `plot(1)` is
    saved. For `fig.keep = 'last'`, the last plot with 10 vertical lines is
    saved.

-   `fig.show`: (`'asis'`; character) How to show/arrange the plots. Possible
    values are:

    -   `asis`: Show plots exactly in places where they were generated (as if
        the code were run in an R terminal);
    -   `hold`: Hold all plots and output them at the end of a code chunk;
    -   `animate`: Concatenate all plots into an animation if there are mutiple
        plots in a chunk;
    -   `hide`: Generate plot files but hide them in the output document.

-   `dev`: (`'pdf'` for LaTeX output and `'png'` for HTML/Markdown; character)
    The graphical device to generate plot files. All graphics devices in base R
    and those in **Cairo**, **cairoDevice**, **svglite**, **ragg**, and
    **tikzDevice** are supported, e.g., `pdf`, `png`, `svg`, `jpeg`, `tiff`,
    `cairo_pdf`, `CairoJPEG`, `CairoPNG`, `Cairo_pdf`, `Cairo_png`, `svglite`,
    `ragg_png`, `tikz`, and so on. See `names(knitr:::auto_exts)` for the full
    list. Besides these devices, you can also provide a character string that is
    the name of a function of the form `function(filename, width, height)`. The
    units for the image size are *always* inches (even for bitmap devices, in
    which DPI is used to convert between pixels and inches).

    The chunk options `dev`, `fig.ext`, `fig.width`, `fig.height`, and `dpi` can
    be vectors (shorter ones will be recycled), e.g., `dev = c('pdf', 'png')`
    creates a `PDF` and a `PNG` file for the same plot.

-   `dev.args`: (`NULL`; list) More arguments to be passed to the device, e.g.,
    `dev.args = list(bg = 'yellow', pointsize = 10)` for `dev = 'png'`. This
    option depends on the specific device (see the device documentation). When
    `dev` contains multiple devices, `dev.args` can be a list of lists of
    arguments, and each list of arguments is passed to each individual device,
    e.g.,
    `dev = c('pdf', 'tiff'), dev.args = list(pdf = list(colormodel = 'cmyk', useDingats = TRUE), tiff = list(compression = 'lzw'))`.

-   `fig.ext`: (`NULL`; character) File extension of the figure output. If
    `NULL`, it will be derived from the graphical device; see
    `knitr:::auto_exts` for details.

-   `dpi`: (`72`; numeric) the DPI (dots per inch) for bitmap devices
    (`dpi * inches = pixels`).

-   `fig.width`, `fig.height`: (both are `7`; numeric) Width and height of the
    plot (in inches), to be used in the graphics device.

-   `fig.asp`: (`NULL`; numeric) The aspect ratio of the plot, i.e., the ratio
    of height/width. When `fig.asp` is specified, the height of a plot (the
    chunk option `fig.height`) is calculated from `fig.width * fig.asp`.

-   `fig.dim`: (`NULL`; numeric) A numeric vector of length 2 to provide
    `fig.width` and `fig.height`, e.g., `fig.dim = c(5, 7)` is a shorthand of
    `fig.width = 5, fig.height = 7`. If both `fig.asp` and `fig.dim` are
    provided, `fig.asp` will be ignored (with a warning).

-   `out.width`, `out.height`: (`NULL`; character) Width and height of the plot
    in the output document, which can be different with its physical `fig.width`
    and `fig.height`, i.e., plots can be scaled in the output document.
    Depending on the output format, these two options can take special values.
    For example, for LaTeX output, they can be `.8\\linewidth`, `3in`, or `8cm`;
    for HTML, they may be `300px`. For `.Rnw` documents, the default value for
    `out.width` will be changed to `\\maxwidth`, which is defined
    [here](../demo/framed/). It can also be a percentage, e.g., `'40%'` will be
    translated to `0.4\linewidth` when the output format is LaTeX.

-   `out.extra`: (`NULL`; character) Extra options for figures. It can be an
    arbitrary string, to be inserted in `\includegraphics[]` in LaTeX output
    (e.g., `out.extra = 'angle=90'` to rotate the figure by 90 degrees), or
    `<img />` in HTML output (e.g.,
    `out.extra = 'style="border:5px solid orange;"'`).

-   `fig.retina`: (`1`; numeric) This option only applies to HTML output. For
    [Retina displays](http://en.wikipedia.org/wiki/Retina_Display), setting this
    option to a ratio (usually 2) will change the chunk option `dpi` to
    `dpi * fig.retina`, and `out.width` to `fig.width * dpi / fig.retina`
    internally. For example, the physical size of an image is doubled, and its
    display size is halved when `fig.retina = 2`.

-   `resize.width`, `resize.height`: (`NULL`; character) The width and height to
    be used in `\resizebox{}{}` in LaTeX output. These two options are not
    needed unless you want to resize TikZ graphics, because there is no natural
    way to do it. However, according to **tikzDevice** authors, TikZ graphics
    are not meant to be resized, to maintain consistency in style with other
    text in LaTeX. If only one of them is `NULL`, `!` will be used (read the
    documentation of **graphicx** if you do not understand this).

-   `fig.align`: (`'default'`; character) Alignment of figures in the output
    document. Possible values are `default`, `left`, `right`, and `center`. The
    default is not to make any alignment adjustments.

-   `fig.link`: (`NULL`; character) A link to be added onto the figure.

-   `fig.env`: (`'figure'`; character) The LaTeX environment for figures, e.g.,
    you may set `fig.env = 'marginfigure'` to get `\begin{marginfigure}`. This
    option requires `fig.cap` be specified.

-   `fig.cap`: (`NULL`; character) A figure caption.

-   `fig.scap`: (`NULL`; character) A short caption. This option is only
    meaningful to LaTeX output. A short caption is inserted in `\caption[]`, and
    usually displayed in the "List of Figures" of a PDF document.

-   `fig.lp`: (`'fig:'`; character) A label prefix for the figure label to be
    inserted in `\label{}`. The actual label is made by concatenating this
    prefix and the chunk label, e.g., the figure label for
    ```` ```{r, foo-plot} ```` will be `fig:foo-plot` by default.

-   `fig.pos`: (`''`; character) A character string for the figure position
    arrangement to be used in `\begin{figure}[]`.

-   `fig.subcap`: (`NULL`) Captions for subfigures. When there are multiple
    plots in a chunk, and neither `fig.subcap` nor `fig.cap` is `NULL`,
    `\subfloat{}` will be used for individual plots (you need to add
    `\usepackage{subfig}` in the preamble). See
    [067-graphics-options.Rnw](https://github.com/yihui/knitr-examples/blob/master/067-graphics-options.Rnw)
    for an example.

-   `fig.ncol`: (`NULL`; integer) The number of columns of subfigures; see
    [here](https://github.com/yihui/knitr/issues/1327#issuecomment-346242532)
    for examples (note that `fig.ncol` and `fig.sep` only work for LaTeX
    output).

-   `fig.sep`: (`NULL`; character) A character vector of separators to be
    inserted among subfigures. When `fig.ncol` is specified, `fig.sep` defaults
    to a character vector of which every N-th element is `\newline` (where `N`
    is the number of columns), e.g., `fig.ncol = 2` means
    `fig.sep = c('', '', '\\newline', '', '', '\\newline', '', ...)`.

-   `fig.process`: (`NULL`; function) A function to post-process figure files.
    It should take the path of a figure file, and return the (new) path of the
    figure to be inserted in the output. If the function contains `options`
    argument, the list of chunk options will be passed to this argument.

-   `fig.showtext`: (`NULL`; logical) If `TRUE`, call
    `showtext::showtext_begin()` before drawing plots. See the documentation of
    the [**showtext**](http://cran.rstudio.com/package=showtext) package for
    details.

-   `external`: (`TRUE`; logical) Whether to externalize tikz graphics
    (pre-compile tikz graphics to PDF). It is only used for the `tikz()` device
    in the **tikzDevice** package (i.e., when `dev='tikz'`), and it can save
    time for LaTeX compilation.

-   `sanitize`: (`FALSE`; character) Whether to sanitize tikz graphics (escape
    special LaTeX characters). See the documentation of the **tikzDevice**
    package.

There are two hidden options that are not designed to be set by the users:
`fig.cur` (the current figure number or index when there are multiple figures),
and `fig.num` (the total number of figures in a chunk). The purpose of these two
options is to help **knitr** deal with the filenames of multiple figures as well
as animations. In some cases, we can make use of them to write animations into
the output using plot files that are saved manually (see the [graphics
manual](https://github.com/yihui/knitr/releases/download/doc/knitr-graphics.pdf)
for examples).

### Animation

-   `interval`: (`1`; numeric) Time interval (number of seconds) between
    animation frames.

-   `animation.hook`: (`knitr::hook_ffmpeg_html`; function or character) A hook
    function to create animations in HTML output; the default hook uses FFmpeg
    to convert images to a WebM video.

    -   Another hook function is `knitr::hook_gifski` based on the
        [**gifski**](https://cran.r-project.org/package=gifski) package to
        create GIF animations.

    -   This option can also take a character string `'ffmpeg'` or `'gifski'` as
        a shorthand of the corresponding hook function, e.g.,
        `animation.hook = 'gifski'` means `animation.hook = knitr::hook_gifski`.

    -   `aniopts`: (`'controls,loop'`; character) Extra options for animations;
        see the documentation of the LaTeX [**animate**
        package](http://www.ctan.org/tex-archive/macros/latex/contrib/animate).

-   `ffmpeg.bitrate` (`1M`; character) To be passed to the `-b:v` argument of
    FFmpeg to control the quality of WebM videos.

-   `ffmpeg.format` (`webm`; character) The video format of FFmpeg, i.e., the
    filename extension of the video.

### Code chunk

-   `code`: (`NULL`; character) If provided, it will override the code in the
    current chunk. This allows us to programmatically insert code into the
    current chunk. For example, `code = readLines('test.R')` will use the
    content of the file `test.R` as the code for the current chunk.

-   `ref.label`: (`NULL`; character) A character vector of labels of the chunks
    from which the code of the current chunk is inherited (see the demo for
    [chunk references](../demo/reference/)).

### Child documents

-   `child`: (`NULL`; character) A character vector of paths of child documents
    to be knitted and input into the main document.

### Language engines

-   `engine`: (`'R'`; character) The language name of the code chunk. Possible
    values can be found in `names(knitr::knit_engines$get())`, e.g., `python`,
    `sql`, `julia`, `bash`, and `c`, etc. The object `knitr::knit_engines` can
    be used to set up engines for other languages.

-   `engine.path`: (`NULL`; character) The path to the executable of the
    `engine`. This option makes it possible to use alternative executables in
    your system, e.g., the default `python` may be at `/usr/bin/python`, and you
    may set `engine.path = '~/anaconda/bin/python'` to use a different version
    of Python.

    `engine.path` can also be a list of paths, which makes it possible to set
    different engine paths for different engines, e.g.,

    ``` r
    knitr::opts_chunk$set(engine.path = list(
      python = '~/anaconda/bin/python',
      ruby = '/usr/local/bin/ruby'
    ))
    ```

    The names of the list correspond to the names of the engines.

### Option templates

-   `opts.label`: (`NULL`; character) The label of options set in
    `knitr::opts_template` (see `?knitr::opts_template`). This option can save
    some typing effort for sets of frequently used chunk options.

### Extracting source code

-   `purl`: (`TRUE`; logical) When running `knitr::purl()` to extract source
    code from a source document, whether to include or exclude a certain code
    chunk.

### Other chunk options

-   `R.options`: (`NULL`; list) Local R options for a code chunk. These options
    are set temporarily via `options()` before the code chunk, and restored
    after the chunk.

## Package-Level Options

The package options can be changed using the object
[`knitr::opts_knit`](/knitr/objects/) (*not to be confused with
`knitr::opts_chunk`*). For example:

``` r
knitr::opts_knit$set(progress = TRUE, verbose = TRUE)
```

See `?knitr::opts_knit` for the alternative approach to setting package options
using the R base function `options()`.

Available package options are:

-   `aliases`: (`NULL`; character) A named character vector to specify the
    aliases of chunk options, e.g., `c(h = 'fig.height', w = 'fig.width')` tells
    **knitr** that the chunk option `h` really means `fig.height`, and `w` is an
    alias for `fig.width`. This option can be used to save some typing effort
    for long option names.

-   `base.dir`: (`NULL`; character) An absolute directory under which the plots
    are generated.

-   `base.url`: (`NULL`; character) The base URL of images on HTML pages.

-   `concordance`: (`FALSE`; logical) Whether to write a concordance file to map
    the output line numbers to the input line numbers. This enables one to
    navigate from the output to the input, and can be helpful especially when
    TeX error occurs. This feature is only for `.Rnw` documents, and implemented
    in RStudio.

-   `eval.after`: (`'fig.cap'`; character) A character vector of option names.
    These options will be evaluated *after* a chunk has been evaluated, and all
    other options will be evaluated before a chunk. For example, for
    `fig.cap = paste('p-value is', t.test(x)$p.value)`, it will be evaluated
    after the chunk according to the value of `x` if `eval.after = 'fig.cap'`.

-   `global.par`: (`FALSE`; logical) If `TRUE`, the `par()` settings from the
    previous code chunk will be preserved and applied to the next code chunk (of
    course, this only applies to base R graphics). By default, **knitr** opens a
    new graphical device to record plots and close it after evaluating the code,
    so `par()` settings will be discarded.

-   `header`: (`NULL`; character) The text to be inserted into the output
    document before the document begins (e.g., after `\documentclass{article}`
    in LaTeX, or after `<head>` in HTML). This is useful for defining commands
    and styles in the LaTeX preamble or HTML header. The beginning of document
    is found using the pattern defined in
    `knitr::knit_patterns$get('document.begin')`. This option is only for `.Rnw`
    and `.Rhtml` documents.

-   `latex.options.color`, `latex.options.graphicx` (`NULL`): Options for the
    LaTeX packages **color** and **graphicx**, respectively. These options are
    only for `.Rnw` documents.

-   `out.format`: (`NULL`; character) Possible values are `latex`, `sweave`,
    `html`, `markdown`, and `jekyll`. It will be automatically determined based
    on the input file, and this option will affect the set of hooks to be set
    (see `?knitr::render_latex` for example). Note this option has to be set
    *before* `knitr::knit()` runs (it will not work if you set it inside the
    document).

-   `progress`: (`TRUE`; logical) Whether to display a progress bar when running
    `knitr::knit()`.

-   `root.dir`: (`NULL`; character) The root directory when evaluating code
    chunks. If `NULL`, the directory of the input document will be used.

-   `self.contained`: (`TRUE`; logical) Whether the output document should be
    self-contained (TeX styles to be written in the `.tex` document, and CSS
    styles to be written in the `.html` document). This option only applies to
    `.Rnw` and `.Rhtml` documents.

-   `unnamed.chunk.label`: (`unnamed-chunk`; character) The label prefix for
    unnamed chunks.

-   `upload.fun`: (`identity`; function) A function that takes a file path,
    processes the file, and returns a character string when the output format is
    HTML or Markdown. Typically, it is a function to upload an image and return
    the link to the image, e.g.,
    `knitr::opts_knit$set(upload.fun = knitr::imgur_upload)` can upload a file
    to <http://imgur.com> (see `?knitr::imgur_upload`).

-   `verbose`: (`FALSE`; logical) Whether to show verbose information (e.g., R
    code in each chunk and message logs), or only show chunk labels and options.
