# R Markdown Conceptual Overview {#conceptual-overview}

The goal of this text is to showcase many tips and tracks for getting the most value from RMarkdown. The following chapters demonstrate techniques to write more efficient and succinct code and to customize your output. Before we begin, it may be helpful to understand just a bit more about how RMarkdown works to help you to understand, remember, apply, and 'remix' these tricks. In this section, we provide a brief overview of the process of knitting a document and the key "levers to pull" to change the output. This material is not necessary to understand the subsequent chapters (so feel free to skip ahead!), but it may help you to build a richer mental model for how all the pieces fit together.

## What happens when we render? {#when-we-render}

**TODO: Could we put Ch2.1 here? Or, if this context material ends up in an Appendix, we can reference 2.1**

## R Markdown Anatomy {#rmd-anatomy}

We can dig one level deeper by considering the different components of an RMarkdown. Specifically, let's look at when and how these are altered during the rendering workflow. 

### YAML Header

The YAML header exercises a wide range of impact on the final document we produce. It can contain many types of information including aspect that effect the code, content, and the rendering process. A typical YAML header might look something like this:

````md
```
title: My R Markdown report
author: Yihui Xie
output: html_document
```
````

In this case, the `title` and `author` fields will be added to the beginning of the report using the default templates, and the `output` field tells RMarkdown to apply the function `rmarkdown::html_document()` when rendering.

We can further influence the rendering processes by passing arguments to the rendering function we are specifying in `output`. For example, writing:

````md
```
output:
  html_document:
    toc: true
    toc_float: true
```
````

is the equivalent of telling `render` to apply the `rmarkdown::html_document()` with the arguments to `toc` and `toc_float` set to `TRUE`. (To find out what these do and to learn about other parameter options, you may run `?rmarkdown::html_document` in your console.)

The YAML header can also influence our content and code itself if we chose to use code parameters, as described in Section \@ref(parameterized-reports). In short, we can include variables and R expressions in this header that can be referenced throughout our RMarkdown. For example, the following header defines `start_date` and `end_date` parameters. `knitr` will use this information to create a list called `params`. Thus, if we want to use these values in our R code, we can access them via `params$x` and `params$y`.

````md
```
title: My RMarkdown
author: Yihui Xie
output: html_document
params:
  start_date: '2020-01-01'
  end_date: '2020-06-01'
```
````

### Narrative

The narrative textual elements of our RMarkdown may be one of the simplest to understand. Typically, this will feel quite a bit like writing in a text editor. However, this markdown content can be more powerful and interesting than simple text -- both in how its content is made and how the document structure is made from it.

While much of our narrative is human-written, many RMarkdown documents will likely wish to reference the code and analysis being used. For this reason, Chapter 3 (**TODO: How to reference without label?**) demonstrates the many ways that code can help generate parts of the text. This conversion is handled by `knitr` as we convert from `.Rmd` to `.md`.

Our markdown text can also provide structure to the document. While we do not have enough space here to review markdown syntax^[Instead, for a review of markdown syntax, please see the [Bookdown book](https://bookdown.org/yihui/bookdown/markdown-syntax.html)] (**TODO: O should we?**), one particularly relevant concept is section headers which are denoted by one or more `#`s (corresponding to different levels).

```
# First-level header

## Second-level header

### Third-level header
```

These headers give structure to our entire document as `rmarkdown` converts the `.md` to our final output format. We can leverage this structure in a number of ways by making reference to these various sections and appending certain formatting and choices to them. We can do this by giving them a unique name by following the header title with `{#name}` or by attaching one or more classes (or 'groups') to them with `{.type}`.

For example, we might write:

```
## Second-level header {#introduction .colorful-text}
```

We can them access this section with many of the tools you'll learn by referencing it by name or by applying options to all sections that are tagged with the same class. As examples, Section \@ref(cross-ref) demonstrates how to use section names to make cross-references throughout your document, and Section \@ref(html-tabs) introduces the `.tabset` class to help reorganize subheaders.

The third and final interesting type of content that we might find in the textual part of our RMarkdown is raw code written in our desired output format (e.g. HTML, LaTeX). This is ignored by both `knitr` and `rmarkdown` and can 'flow through' to be correctly interpreted in the final output.

### Code Chunks

Code chunks are the beating heart of our RMarkdown. The code in these chunks is run by `knitr` and its output is translated to markdown to dynamically keep our reports in-sync with our current scripts and data. Each code chunk consists of a language engine (\@ref(other-languages)), an optional name, chunk options (\@ref(chunk-options)), and code. 

To understand some of the modifications we can make to code chunks, it is worth understanding the `knitr` process in slightly more detail. For each chunk, a `knitr` language engine receives as input the knitting environment, the code input, a set of chunk options, and returns that environment, a formatted representation of the code, and a formatted representation of the output.

```{r knitr-workflow, echo = FALSE, fig.cap = 'A flowchart of inputs and outputs to a language engine.', out.width = '50%'}
nomnoml::nomnoml(
  "
  [<frame>Code chunk|
  [Code]->[Language Engine]
  [Chunk Options]->[Language Engine]
  [Environment]->[Language Engine]
  [Language Engine]->[Formatted Code]
  [Language Engine]->[Formatted Output]
  [Language Engine]->[Enviornment]
  ]")
```

We can modify this process by changing our language engine; modifying chunk options which can be global, local, or engine-specific; and by using hooks (Chapter \@ref(chunk-hooks) and \@ref(output-hooks)) to further process the these inputs and outputs. For example, in Section \@(hook-hide), you'll learn how to add a hook to post-process the code output (the 'source') to refact certain lines.

Code chunks also have analogous concepts to the groups and unique identifiers that we explored for narrative sections. A code chunk can specify an optional name immediately after its language engine, and it can set classes for the code chunk our the output by using the `class.source` and `class.output` code chunk options. For example, the chunk header `{r summary-stats, class.output = 'bigplot'}`
 gives this chunk the unique name `summary-stats` and the class of `bigplot`. A chunk can have only one name but can have multiple classes.

### Document Body

One important thing to understand when authoring and modifying a document is how code and narrative pieces create different sections, or containers within the document. For example, suppose we have a document that looks something like this:

````md
# Title

## Section x

This is my introduction

```{r chunk-x}`r ''`
x <- 1
print(x)
```
### Subsection 1

Here are some details

### Subsection 2

These are more details

## Section y

This is another section

```{r chunk-y}`r ''`
y <- 2
print(y)
```
````

When writing this document, we likely think of each piece as linear with independent sections of text and code following in a sequence one after the other. However, what we are actually doing is creating a set of nested containers which conceptually^[In reality, there are many more containers than shown. For example, for a knitted code chunk, the code and output exist in separate containers which share a common parent] looks more like this: 

```{r rmd-containers, echo = FALSE, fig.cap = 'A simple RMarkdown illustrated as a set of nested containers.', out.width = '50%'}
nomnoml::nomnoml(
  "
  [Title (Level 1)|
  
    [Section x (Level 2)| - Text | - Code (chunk-x) | - Subsection 1 | - Subsection 2]
    [Section y (Level 2)| - Text | - Code (chunk-y) ]

  ]")
```

Two key features of this diagram are that every section of text or code is it's own discrete container and that containers can be nested within one another. This nesting is particularly apparent if you are authoring your RMarkdown in the RStudio IDE and expand the document outline.

Note that in the image about, header of the same level (the same number of `#`'s) represent containers at the same level of nesting. More minor headers (those with a larger number of `#`'s than the preceding) exist inside of the container of more major headers (those with smaller number of `#`s). In such a case, it is common to call the major sections the 'parent' and the minor sections the 'child'. Beyond the use of headers, you can also create divisions in your code using `:::`, as demonstrated in **!!TODO: can't reference the multicolumn layout section because it doesn't have an ID!!**. 

This structure has important implications as we attempt to apply some of the formatting and styling options that are described in this text. For example, we will see this nested structure when we learn about how Pandoc represents our document as an abstract syntax tree (Section \@ref(lua-filters)) or when we use CSS selectors (Section \@ref(html-css), among others) to style our HTML output. 

Some options we can chose to apply to containers of similar types (e.g. code section) and others we can chose to apply to all containers that exist inside of another container (e.g. everything under 'Section y'). Additionally, as explained above, we can use classes designate some sections as being more or less similar by adding attributes to denote their common properties or intent. 

As you read through this cookbook, it may be useful to quiz yourself and think about what sort of container the specific 'recipe' is acting upon.

## What can we change to change the results? {#what-to-change}

Let's summarize what we've seen so far and preview what's to come.

Rendering an output from RMarkdown consists of converting `.Rmd` to `.md` with `knitr` and then `.md` to our desired output with `pandoc` (typically). 

The `.Rmd` to `.md` step handles the execution and "translation" of all code within our report, so most changes to *content* involve editing the `.Rmd` with code for `knitr` to translate. Some tools we have to control these steps are chunk options and hooks.

Our `.md` is a plain text file with no formatting. This is where `pandoc` comes in to convert to HTML, PDF, or Word. Along the way, we add structure and style. A wide range of tools help us in this process including style sheets, raw LaTeX or HTML code, pandoc templates, and Lua filters.  By understanding the nested structure of an RMarkdown and by thoughtfully using identifiers and classes, we can apply some of these tools selectively to targetted parts of our output.

Finally, our YAML header may help us toggle any of these steps. Changing parameters can change how our code runs, changing metadata alters the text content, and changing output options provides the render function with a different set of instructions.

Of course, these are all rough heuristics and should not be taken as absolute fact. Ultimately, there is not a completely clear division of labor. Throughout this book, you will see that there are often multiple valid paths to achieving many of the outcomes described in this book, and these may enter different stages of the pipeline. For example, for the simple task of inserting in image in your document, you may use the R code  `knitr::include_graphics()` which would execute in the `.Rmd` to `.md` stage or you could directly use markdown syntax (`![]()`). This may seem confusing, and sometimes different approaches will have different advantages. However, do not be concerned; if anything, this means you often means there are many valid ways to solve your problem and you can follow whichever approach makes the most sense to you.

And that's that! In the rest of the book, you can now color in this rough sketch with many more concrete examples of ways to modify all of the components we've discussed to get the most value out of RMarkdown.
  
