--- 
title: "R Markdown Cookbook"
author: "Michael Harper, Yihui Xie"
date: "`r Sys.Date()`"
description: Tips and tricks to build advanced projects in R Markdown
site: bookdown::bookdown_site
documentclass: book
bibliography: [book.bib, packages.bib]
biblio-style: apalike
link-citations: yes
---

# About the book

> This book is in a very early stage of development. If you have any suggestions on what should be included within this book, please get in touch via GitHub: https://github.com/mikey-harper/rmarkdown-cookbook

```{r include=FALSE}
install.packages("DiagrammeR")
install.packages("webshot")
webshot::install_phantomjs()

library(webshot)
library(DiagrammeR)
```



- This book aims to bring together lots of useful tips for R Markdown.
- One of the common criticisms of markdown as a language is that it naturally limits what you can write. Users often therefore feel limited in what they can achieve with advanced customisation. However, we can easily edit.

It is noted in the book *R Markdown: The Definitive Guide* that "Please do not underestimate the customizability of R Markdown because of the simplicity of its syntax"

## Recommended Reading

There are already several fantastic books out there which you may have already read:

- Dynamic documents and knitr
- R Markdown: The Definitive guide
- Authoring books with bookdown
- Blogdown


```{r include=FALSE}
# automatically create a bib database for R packages
knitr::write_bib(c(
  .packages(), 'bookdown', 'knitr', 'rmarkdown', 'DiagrammeR'
), 'packages.bib')
```

