--- 
title: "R Markdown Cookbook"
author: "Michael Harper, Yihui Xie"
date: "`r Sys.Date()`"
description: Tips and tricks to build advanced projects in R Markdown
site: bookdown::bookdown_site
documentclass: book
bibliography: [book.bib, packages.bib]
biblio-style: apalike
link-citations: yes
---

> This book is in a **very** early stage of development. If you have any suggestions on what should be included within this book, please get in touch via [GitHub](https://github.com/mikey-harper/rmarkdown-cookbook)

```{r include=FALSE}
install.packages("DiagrammeR")
install.packages("webshot")
webshot::install_phantomjs()

library(webshot)
library(DiagrammeR)
```

# Preface {-}

- One of the common criticisms of markdown as a language is that it naturally limits what you can write. Users often therefore feel limited in what they can achieve with advanced customisation. However, there are a growing range of options available to achieve tailor the outputs of documents.
- This book aims to bring together useful tips for R Markdown.
- Although there are many resources available, many of these cover the basics of R Markdown.

## Recommended Reading {-}

There are already several fantastic books out there which you may have already read:

- Dynamic documents and knitr
- R Markdown: The Definitive guide
- Authoring books with bookdown
- Blogdown

```{r include=FALSE}
# automatically create a bib database for R packages
knitr::write_bib(c(
  .packages(), 'bookdown', 'knitr', 'rmarkdown', 'DiagrammeR'
), 'packages.bib')
```

