## Controlling the Placement of Figures
<!--- https://stackoverflow.com/questions/16626462/figure-position-in-markdown-when-converting-to-pdf-with-knitr-and-pandoc/17648350#17648350 --->
<!--- Some of the solutions adapted from https://texfaq.org/FAQ-floats. Link left here for future reference --->

One of the common frustrations with LaTeX is how Figures and Tables are handled. Unlike in a word processor like *Microsoft Word*, where figures are placed directly where the user specifies, LaTeX will attempt to place the figure in a position which does not violate certain typographic rules. In doing so, it can result in figures being located away from where they are referenced in the text. This section will explain some background information on how floats work, before providing several options for how to customise their behaviour for your work.

### Floats

A detailed description of floats is available within the [LaTeX Wiki](https://en.wikibooks.org/wiki/LaTeX/Floats,_Figures_and_Captions), but in summary, floats are used as containers for things which cannot be broken over a page, such as tables and figures.  By default, if the figure or table can not be contained in the space left in the current page, it is placed at the top of the next page. This behaviour can be controlled by different placement specifiers as follows:

- `h`:	Place the float here, i.e., approximately at the same point it occurs in the source text (however, not exactly at the spot)
- `t`:	Position at the top of the page.
- `b`:	Position at the bottom of the page.
- `p`:	Put on a special page for floats only.
- `!`:	Override internal parameters LaTeX uses for determining "good" float positions.
- `H`: Places the float at precisely the location in the LaTeX code. Requires the float package (`\usepackage{float}`)

The default behaviour for LaTeX in R Markdown is `tbp`, that is it will try positioning the figure at the top of the page, then then bottom, and if not it will put them on a separate page for floats.

### Preventing figures floating

Many users will initially want to prevent figures from floating in their document, replicating the behaviour of a traditional word processor. To do this, we must firstly load the LaTeX package `float`. This can be done by including the following line in the YAML:

```
output: 
pdf_document:
extra_dependencies: ["float"]
```

We can use the `fig.pos` to control the float behaviour, and the use of `!H` will prevent any floating within the document. We can set the default behaviour for the document so that all chunks have this setting by including the following line at the start of your R Markdown document:

```{r, echo=TRUE, eval=FALSE}
knitr::opts_chunk$set(fig.pos = "!H")
```

Please note, that we would strongly advise that readers consider the behaviour of float in their document. This solution was included within the book by popular demand^[the related[StackOverflow question](https://stackoverflow.com/q/16626462/7347699) has been viewed over 40000 times] but there are some serious sides effects overriding the default settings. Primarily, you may end 

### Force floats forwards
<!--- https://tex.stackexchange.com/questions/15706/force-floats-to-be-typeset-after-their-occurrence-in-the-source-text --->

An alternative to forcing all floats to be held is to force floats forwards in the text. This can remove a common issue, where a figure is shown at the top of the page on which it is created. This can break the flow of a report. We can force it so that the figure always appears after the text by using the **flafter** LaTeX package as follows:

```
output: 
pdf_document:
extra_dependencies: ["flafter"]
```

### Adjust LaTeX placement rules

LaTeX’s own float placement parameters could be preventing placements that seem entirely “reasonable” to you — they’re notoriously rather conservative. These defaults are displayed in Table \@ref(tab:latexFloatDefault).

<!--- This table may work better as a diagram? --->

```{r latexFloatDefault}
data.frame(
  Command = c("topfraction", "bottomfraction", "textfraction",
              "floatpagefraction", "topnumber", "bottomnumber",
              "totalnumber"),
  Description = c("maximum fraction of page for floats at top",
                  "maximum fraction of page for floats at bottom",
                  "minimum fraction of page for text",
                  "minimum fraction of floatpage that should have floats",
                  "maximum number of floats at top of page", "maximum number of floats at bottom of page",
                  "maximum number of floats on a page"),
  Default = c("0.7", "0.3", "0.2", "0.5", "2", "1", "3"),
  stringsAsFactors=FALSE
) %>%
  knitr::kable(caption = "Default LaTeX float settings", escape = TRUE)

```

To encourage LaTeX not to move your figure, we can alter these default settings. We could include the following in our LaTeX preamble file, reducing the minimum amount of text required on a page and allow more room for floats:

```latex
\renewcommand{\topfraction}{.85}
\renewcommand{\bottomfraction}{.7}
\renewcommand{\textfraction}{.15}
\renewcommand{\floatpagefraction}{.66}
\renewcommand{\dbltopfraction}{.66}
\renewcommand{\dblfloatpagefraction}{.66}
\setcounter{topnumber}{9}
\setcounter{bottomnumber}{9}
\setcounter{totalnumber}{20}
\setcounter{dbltopnumber}{9}
```
If we added these changes to our `premable.tex` file, we could reference to them within our YAML of our document:

```yaml
output:
    pdf_document:
        includes:
            in_header: premable.tex
```


