## Customising kable outputs

The `kable` package provides the most convenient way of producing tables within the R Markdown package, providing neatly formatted tables for PDF, HTML, Word and Markdown documents. The majority of the time, the default settings are generally suitable, but if we wish to have more control of their behaviour, there is an extensive number of settings which can be used. Table \@ref(tab:kableArgs) shows the possible arguments which can be used, and with some examples shown below on how they can be used.

```{r kableArgs, echo = FALSE}
filter_out <- c("x", "...")
to_table <- function(args, HTML = FALSE, Latex = FALSE, Markdown = FALSE) {
  tibble::tibble(
    Argument = names(args),
    Description = NA,
    HTML = HTML,
    Latex = Latex,
    Markdown = Markdown,
    Default = vapply(args, rlang::as_label, FUN.VALUE = "")
  )
}
extract_args <- function(fun) {
  args <- rlang::fn_fmls(fun)
  args <- args[setdiff(names(args), filter_out)]
}

# get all args from default function
all_args <- to_table(extract_args(knitr::kable), HTML = TRUE, Latex = TRUE, Markdown = TRUE)
# get args from specific html function
html_args <- to_table(extract_args(knitr:::kable_html), HTML = TRUE)
html_args <- html_args[!html_args$Argument %in% all_args$Argument,]
# get args from specific latex function
latex_args <- to_table(extract_args(knitr:::kable_latex), Latex = TRUE)
latex_args <- latex_args[!latex_args$Argument %in% all_args$Argument,]
# get args from specific markdown function
markdown_args <- to_table(extract_args(knitr:::kable_markdown), Markdown = TRUE)
markdown_args <- markdown_args[!markdown_args$Argument %in% all_args$Argument,]
# format table
rbind(all_args, html_args, latex_args, markdown_args) %>%
  knitr::kable(caption = "Possible arguments for the kable function")
```

### Changing column names
<!--- https://stackoverflow.com/questions/51432502/replace-column-names-in-kable-r-markdown/51444998#51444998 --->

The names of columns in a dataframe will often not be the same as what we want to display in our output format. It can be particularly annoying to have spaces in the name of a variable. We can use the the `col.names` argument to replace the names with a vector of names:

```{r}
knitr::kable(head(iris),
      col.names = c("Sepal Length", "Sepal Width", "Petal Length", "Petal Width", "Species"))
```

### Changing column alignment

If we wish to change the alignment of a knitr table, we can use either a vector of values or a single multi-character string for alignment, so `kable(..., align=c('c','l'))` can be shortened to `kable(..., align='cl')`.

```{r}
knitr::kable(head(mtcars), align = "llrrcc")
```


The shortest form can't be used with _latex_ format. You must use longer form using a vector
```{r, eval = FALSE}
knitr::kable(head(mtcars), align = c("l", "l", "r", "r", "c", "c"), format = "latex")
```


### Customising LaTeX options

There are a lot of extra options which can be controlled for LaTeX tables. 

- `booktabs` [booktabs](https://ctan.org/pkg/booktabs?lang=en) LaTeX package.
- If we are creating tables which are longer than a page, we can  `longtable` option. This will load the [longtable](https://ctan.org/pkg/longtable?lang=en) LaTeX package. 
- The `caption.short` argument can be useful if we wish to show a shortened caption in our table of contents.
- `linesep` can be used to change the line spacing. We can play around with this option if you want to  change the style. For example `linesep = c("", "", "", "\\hline")` would add a horizontal line every four spaces.

### Customising HTML table appearance
<!--- https://stackoverflow.com/questions/24254552/knitr-style-table-with-css --->

### Changing default behavour

There are many global options which apply to kable which can be set within your R Markdown document. These can be controlled globally within your R Markdown document using the `options` function. The settings which can be controlled include:

- `options(knitr.kable.NA = '')`
- ADD MORE


