---
title: Show chunk headers in the output
---

Set up a chunk hook named `wrapper` to wrap the chunk
output inside the original chunk header and footer.

```{r, setup, include=FALSE}
knitr::knit_hooks$set(wrapper = function(before, options) {
  # the original chunk might be indented
  if (is.null(indent <- options$indent)) indent <- ''
  
  # hide the wrapper=TRUE option
  opts <- gsub(', wrapper=TRUE', '', options$params.src)
  
  if (before) {
    # add the header
    sprintf('\n\n%s````\n```{r,%s}\n````\n', indent, opts)
  } else {
    # add the footer
    sprintf('\n\n%s````\n```\n````\n', indent)
  }
})
```

Now we apply the hook via the chunk option `wrapper=TRUE`.
Remember to put `wrapper=TRUE` at the end of the header, and
it has to be `wrapper=TRUE` precisely (e.g., not `wrapper=T`),
following a comma and a space, unless you adjust the `gsub()`
call in the above hook.

```{r, test-label, collapse=TRUE, wrapper=TRUE}
1 + 1
plot(cars)
```

You should see the original chunk header appear in
the output. The hook should also work when the chunk
is indented, e.g.,

- One bullet.

  ```{r, eval=TRUE, wrapper=TRUE}
  2 + 2
  ```

- Another bullet.
