---
title: Figure Evaluate Caption After
output: bookdown::pdf_document2
---

Knitr can automatically generate a figure environment if fig.cap is specified.

```{r test, fig.cap= paste('$2 \times 2$ is', 2*2), fig.height=2.5}
par(mar=c(4, 4, .1, .1)); plot(runif(4))
```

# Eval After

When `fig.cap` needs to use objects in the current chunk, we need to set the eval.after options so the objects are available when `fig.cap` is used:

```{r}
library(knitr)
# evaluate fig.cap after a chunk is evaluated
opts_knit$set(eval.after = 'fig.cap')
```

Dynamically generate a figure caption:

```{r t-test, fig.cap=paste('The P-value is', format.pval(t.test(x)$p.value)), fig.height=2.5}
x <- rnorm(30)
par(mar=c(4, 4, .1, .1)); hist(x, main='')
```

Change the figure environment:

```{r figure-asterisk, fig.env='figure*', fig.cap='Figure * environment.', fig.width=2, fig.height=4}
par(mar=c(4, 4, .1, .1)); boxplot(x)
```

If you have two figures, you can assign two captions to fig.cap:

```{r two-caps, fig.cap=c('Hi figure one', 'Hi figure two'), fig.width=2, fig.height=1.5, fig.align='center'}
par(mar=rep(0, 4), ann=FALSE)
plot(1:10)
plot(runif(10))
```

You can use figure references too. For example, Figure \ref{fig:t-test} is a histogram, and Figure \ref{fig:figure-asterisk} is a boxplot. The reference key is ``fig: + the chunk label'' by default.
