---
title: Output figures in `<figure>` tags
output: html_document
---

Given a plot file path `x` and a figure caption in the chunk
option `options$fig.cap`, we want to write the plot in the
HTML5 tag in this form:

```html
<figure>
  <img src="PATH" alt="CAPTION" />
  <figcaption>CAPTION</figcaption>
</figure>
```

Now we redefine the `plot` hook (only when the output format
is HTML):

```{r}
if (knitr::is_html_output()) knitr::knit_hooks$set(
  plot = function(x, options) {
    cap  <- options$fig.cap  # figure caption
    tags <- htmltools::tags
    as.character(tags$figure(
      tags$img(src = x, alt = cap),
      tags$figcaption(cap)
    ))
  }
)
```

The plot from the code chunk below will be placed in the
`<figure>` tag:

```{r, fig.cap='A scatterplot for the cars data.'}
par(mar = c(4.5, 4.5, .2, .2))
plot(cars, pch = 19, col = 'red')
```

We add some CSS styles to "see" the `<figure>` and
`<figcaption>` tags better (the `figure` has a dashed
border, and the caption has a light pink background):

```{css, echo=FALSE}
figure {
  border: 2px dashed red;
  margin: 1em 0;
}
figcaption {
  padding: .5em;
  background: lightpink;
  font-size: 1.3em;
  font-variant: small-caps;
}
```

