---
title: Move Code out of the Figure and Table Environments
author: Yihui Xie
output: bookdown::pdf_document2
documentclass: article
---

Sometimes we may not want the code to be shown inside a float, such as a figure or table environment. The hook function `knitr::hook_movecode()` was designed for this purpose; it serves a document hook to post-process the LaTeX results.

First we set up the document hook:

```{r}
library(knitr)
knit_hooks$set(document = hook_movecode)
```

Now we test a figure. You will see there is only a plot in Figure \@ref(fig:example-a); the code was moved out of the environment, and shown
below right after this paragraph.

```{r example-a, fig.height=4, fig.cap = "A figure caption here, and you must label it!"}
summary(cars)
par(mar=c(4,4,.1,.1))
plot(cars)
```

Similarly, you can work with tables; see Table \ref{tab:example-b}.

```{r example-b,  results='asis'}
kable(head(iris),
      caption = "A table caption here; again, you must label it.")
```

If you do not label the figures or tables, or they do not have captions, the hook function `hook_movecode()` will not touch them.


