# (PART) Additional Packages {-} 


# Diagrams

It is often useful to express analysis using flowcharts and diagrams within reports. Although there are many separate programs which can be used to produce these, it can be beneficial to create these within our analysis directly. This makes it easier to update and edit the graph if we need to edit it in the future, and allows us to use the results from our analysis directly within the diagrams, making them more information.

While there are several different packages available for R, we will focus on the package **DiagrammeR** [@R-DiagrammeR]. We also recommend that you read at http://rich-iannone.github.io/DiagrammeR/index.html . In this chapter we will explain some of the basic usages of the package.

## Basic Diagrams

**DiagrammeR** provides methods to build graphs for a number of different graphing languages.

R Studio provides native support for Graphviz (`.gv`) and mermaid (`.mmd`) files. Using files of these types in RStudio provides the advantage of syntax coloring and allowing a quick preview of the diagram. For example, we can make a simple flowchart with the following code:

```{r diagrammeR-basic,  fig.align='center', fig.asp=2, fig.cap="A basic graphic using DiagrammeR", fig.height=2, fig.width=6, out.width="100%"}

DiagrammeR::grViz("digraph {

        graph [layout = dot, rankdir = LR]
        
        node [shape = rectangle]        
        rec1 [label = 'Step 1']
        rec2 [label = 'Step 2']
        rec3 [label =  'Step 3']
        rec4 [label = 'Step 4']

        # edge definitions with the node IDs
        rec1 -> rec2 -> rec3 -> rec4
        }",
        height = 200)
```

There are extensive controls which can be used to control the shape of nodes, colours, line types and add additional parameters.

## Adding parameters to plots

Graphviz substitution allows for mixing in R expressions into a Graphviz graph specification without sacrificing readability. If you specify a subsitution with `@@`, you must ensure there is a valid expression for that substitution. The expressions are placed as footnotes and their evaluations must result in an R vector object (i.e., not a data frame, list, or matrix). Because there is the possibility to have multiple substitutions, numbering is required. Thus, the @@ notation is immediately followed by a number and that number should correspond to the number of the footnoted R expression. For example, suppose we have a dataset which is being analysed, and we would like to identify how many values are removed at each stage of a process.

```{r diagramme-parameters}

DiagrammeR::grViz("
digraph graph2 {

graph [layout = dot, rankdir = LR]

# node definitions with substituted label text
node [shape = rectangle]
a [label = '@@1']
b [label = '@@2']
c [label = '@@3']
d [label = '@@4']

a -> b -> c -> d

}

[1]: paste('Input Data', 1)
[2]: paste('Step ', 1)
[3]: paste('Step ', 1)
[4]: paste('Step ', 1)
")

```

