\mainmatter

# Installation

We assume you have already installed R (https://www.r-project.org) [@R-base] and the RStudio IDE (https://www.rstudio.com). RStudio is not required but recommended, because it makes it easier for an average user to work with R Markdown. If you do not have RStudio IDE installed, you will have to install Pandoc\index{Pandoc} (http://pandoc.org), otherwise there is no need to install Pandoc separately because RStudio has bundled it. Next you can install the **rmarkdown** package in R:

```{r eval=FALSE}
# Install from CRAN
install.packages('rmarkdown')

# Or if you want to test the development version,
# install from GitHub
if (!requireNamespace("devtools"))
  install.packages('devtools')
devtools::install_github('rstudio/rmarkdown')
```

## Installing LaTeX with TinyTeX

If you would like to create PDF outputs from R Markdown, you will need to have a LaTeX distribution installed. Although there are several traditional options including MiKTeX, MacTeX, and TeX Live, these can be cumbersome to install if you primarily want to use LaTeX within R Markdown.

To make this process easier, TinyTeX is a custom LaTeX distribution based on TeX Live that is small in size (150Mb on macOS/Linux and 220Mb on Windows) but functions well in most cases, especially for R users. Installing or running TinyTeX does not require admin privileges, which means you no longer need sudo or your IT.

If you want to use TinyTeX, we recommend that you first uninstall your existing LaTeX distribution. The **tinytex** R package [@R-tinytex] can be used to install TinyTeX:

```{r eval=FALSE}
install.packages('tinytex')
tinytex::install_tinytex()
# to uninstall TinyTeX, run tinytex::uninstall_tinytex() 
```
